---
title: "lfmix"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{lfmix}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# "lfmix: Mixed frequency analysis of low flows using mixed distribution and mixed copula estimators"

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mixdist)
library(lfstat)
library(FAdist)
```

# Introduction of package lfmix
blabla
two estimators. 
- Mixed distribution estimator `weiMixT`
- Mixed copula estimator `weiMicTC`

## 1. Method weiMixT

The method `weiMixT` calculates empirical and theoretical probabilities of annual low flow events.

```{r}
output <- weiMixT(x1=xa, sortx=FALSE, T=c(100, 50, 20))
```

The output is a list with following main elements:

### $sample

A `data.frame` containing event characteristics of annual low flow events with folowing colums: hyear is the hydrological year, x is the annual minimum flow. The `p.` (and `T.`) are event probabilities (and return periods) from various probability estimators: \
**Empirical probabilities (and return periods)** \
`pobs` are empirical probabilities using the classical Weibull estimator \
`pobs.mix` empirical return periods from the mixed distribution approach \
**Theoretical probabilities (and return periods)** \
`pA` (`TA`)... classical anual minima series approach \
`pMix` (`TMix`) ... mixed distribution estimator from Laaha (2023a) \
`pMixC` (`TMixC`) ... mixed copula estimator from Laaha (2023b) \


```{r}
head(output$sample)
```


### $To_events: 

A `data.frame` containing the nominal and improved probabilities (`p.`) and return periods (`T.`) of of ´design´ events with specified return periods from various estimators. The design events are specified by argument `T` with default to 100, 50 and 20 year events.
`To` (`po`) ... nominal values corresponding to classical annual minimum estimator (AMS approach)\
`Tmix` (`pmix`) ... mixed distribution estimator from Laaha (2023a) \
`TmixC` (`pmixC`) ... mixed copula estimator from Laaha (2023b) \
`Ts` (`ps`) ... seasonal estimate from the marginal summer distribution) \
`Tw` (`pw`) ... seasonal estimate from the marginal winter distribution \
```{r}
output$To_events
```

### Various other flow characteristics

`SI`         `SR`           `Rho`          `Rho_p-value`  `mixture_rate`


## 2. Steps of analysis

### Data
```{r}
x1 <- xa
head(x1)
```
### Circular seasonality index (SI)
```{r}
SI <- seasindex(x1)
print(SI)
```
### Seasonality ratio (SR)
```{r}
SR <- seasratio(x1, breakdays=c("01/04", "01/11"))
SR <- as.numeric(SR)
print(SR)
```


### Separate the time series x1 into sum/win season 
```{r}
x1 <- xa
x1s <- x1[ x1$month >= 4 & x1$month <= 10 , ]
x1w <- x1[ x1$month < 4  | x1$month > 10 , ]
```

### Prepare annual and seasonal series
```{r}
AM0 <- MAM(x1,n=7, yearly = TRUE)
AM <- AM0$MAn
lmom <- samlmu(AM)
x.pelwei <- pelwei(lmom); x.pelwei
x.p <- cdfwei(AM, para = x.pelwei)

AM0s <- MAM(x1s,n=7, yearly = TRUE)
#if (sortx==TRUE) {AM0s <- AM0s[order(AM0s$MAn),]}
AMs <- AM0s$MAn
lmoms <- samlmu(AMs)
x.pelwei.s <- pelwei(lmoms); x.pelwei.s
x.ps <- cdfwei(AM, para = x.pelwei.s)

AM0w <- MAM(x1w,n=7, yearly = TRUE)
#if (sortx==TRUE) {AM0w <- AM0w[order(AM0w$MAn),]}
AMw <- AM0w$MAn
lmomw <- samlmu(AMw)
x.pelwei.w <- pelwei(lmomw); x.pelwei.w
x.pw <- cdfwei(AM, para = x.pelwei.w)
```

### Correlation and mixture rate
```{r}
AM0sw <- merge(AM0s, AM0w, by = "hyear", all = TRUE)
cor_spearman <- cor.test(AM0sw$MAn.x, AM0sw$MAn.y, method = "spearman")
print(cor_spearman)
mixture_rate <- sum(AMs < AMw)/length(AM)
print(mixture_rate)
```

### Empirical probabilites (Weibull plotting positions)
```{r}
# Common Weibull prob estimators -> Same as pobs(AM) from package copula! 
.pobs <- (rank(AM))/(length(AM) + 1) # Weibull prob estimators -> Same as pobs(AM) from package copula!
print(.pobs)

# Empirical mixed prob 
.pobs.mix <- pobs_mixed_FUN(AM=AM, AMs=AMs, AMw=AMw, plot=FALSE) # $AM, $p.mix and $p.mix.c are returned
print(.pobs.mix)
```

### Predicted mixed prob (Weibull cdf model)
```{r}
# Mixed distribution approach
x.pred.mix <- cdfwei_Mixed(x=AM, y=AM, para.x = pelwei(samlmu(AMs)), para.y = pelwei(samlmu(AMw)), plot=FALSE)
print(x.pred.mix)

# Predicted mixed prob with copula (model)
x.pred.mixC <- cdfwei_MixedC(x=AM, y=AM, para.x = pelwei(samlmu(AMs)), para.y = pelwei(samlmu(AMw)), x1=x1, plot=FALSE)
print(x.pred.mixC)
```

### Build output object `$sample`
```{r}
#obs <- data.frame(cbind(hyear=AM0$hyear, x=AM, pobs=.pobs, pobs.mix=.pobs.mix$p.mix, pobs.mix=.pobs.mixC$p.mix.c))
sample <- data.frame(hyear=AM0$hyear, x=AM, pobs=.pobs, pobs.mix=.pobs.mix$p.mix, pA=x.p, TA=1/x.p, pMix=x.pred.mix, TMix=1/x.pred.mix, "pMixC"=x.pred.mixC, "TMixC"=1/x.pred.mixC)
```


## 2. Return period (and probability) estimates of design events by various estimators

Here we show how the `$To_events` are calculated. As stated above, the aim is to produce a `data.frame` containing the nominal and improved probabilities (`p.`) and return periods (`T.`) of of ´design´ events with specified return periods from various estimators. The design events are specified by argument `T` with default to 100, 50 and 20 year events.

The `data.frame` will allow us to analyse the change in return period when using our improved estimators 

### Annual Series (no seasons, common probability distribution)
```{r}
T=c(100,50, 20)
.f <- 1/T # Hier Probability 1/T
.y <- quawei(.f, para = x.pelwei) # Hier Quantilwerte von T => als Eingang in andere Prob-Berechnungen
.p <- cdfwei(.y, para = x.pelwei)
pred <- data.frame(yo=.y, po=.p, To=1/.p)
print(pred)
```

### Seasonal distributions
```{r}
# -) summer
.ys <- quawei(.f, para = x.pelwei.s)
.ps <- cdfwei(.y, para = x.pelwei.s)
pred.s <- data.frame(y_of_To=.ys, p=.ps, T=1/.ps)
print(pred.s)
# -) winter
.yw <- quawei(.f, para = x.pelwei.w)
.pw <- cdfwei(.y, para = x.pelwei.w)
pred.w <- data.frame(y_of_To=.yw, p=.pw, T=1/.pw)
print(pred.w)
```

### Mixed distribution estimator
```{r}
.pmix <- cdfwei_Mixed(x=.y, y=.y, para.x = pelwei(samlmu(AMs)), para.y = pelwei(samlmu(AMw)), plot=FALSE, pch="x")

# Ermittle Quantile der Mischverteilung mit p=.f (z.b. p=1/20=0.2)
# Dazu feine Diskretisierung
t.y1 <- par("usr")[3] ; t.y2 <- par("usr")[4] ### wie ohne plot berechnen???
t.y.qua <- seq(round(t.y1, 1), round(t.y2, 1), by=0.01)
t.pmix <- cdfwei_Mixed(x=t.y.qua, y=t.y.qua, para.x = pelwei(samlmu(AMs)), para.y = pelwei(samlmu(AMw)), plot=FALSE, pch="x")
# Nun Suche an der y-Stelle der Verteilung wo x=.pmix=.f ist
# ... d.h, größter Wert mit p <= .f
f.qmix <- vector()
for (i in 1:length(.f)){
  f.qmix[i] <- max(t.y.qua[t.pmix <= .f[i]]) ### better vectorize ...
}

pred.mix <- data.frame(y_of_To=f.qmix, p=.pmix, T=1/.pmix)
print(pred.mix)
```

### Mixed copula estimator
```{r}
.pmixC <- cdfwei_MixedC(x=.y, y=.y, para.x = pelwei(samlmu(AMs)), para.y = pelwei(samlmu(AMw)), x1=x1, plot=FALSE)

# Ermittle Quantile der Mischverteilung mit p=.f (z.b. p=1/20=0.2)
# Dazu feine Diskretisierung
t.y1 <- par("usr")[3] ; t.y2 <- par("usr")[4] ### wie ohne plot berechnen???
t.y.qua <- seq(round(t.y1, 1), round(t.y2, 1), by=0.01)
t.pmixC <- cdfwei_MixedC(x=t.y.qua, y=t.y.qua, para.x = pelwei(samlmu(AMs)), para.y = pelwei(samlmu(AMw)), x1=x1, plot=FALSE, pch="x")
# Nun Suche an der y-Stelle der Verteilung wo x=.pmix=.f ist
# ... d.h, gr??er Wert mit p <= .f
f.qmixC <- vector()
for (i in 1:length(.f)){
  f.qmixC[i] <- max(t.y.qua[t.pmixC <= .f[i]]) ### better vectorize ...
}

pred.mixC <- data.frame(y_of_To=f.qmixC, p=.pmixC, T=1/.pmixC)
print(pred.mixC)
```

### Finally build output object `$To_events`
```{r}
To_events <- data.frame("To"=1/.p, "po"=.p, "Tmix"=1/.pmix , "pmix"=.pmix , "TmixC"=1/.pmixC , "pmixC"=.pmixC , "Ts"=1/.ps, "ps"=.ps, "Tw"=1/.pw, "pw"=.pw)
print(To_events)
```

